---
title: "BFF_SV_analysis"
author: "Lorena Derezanin"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# initialize renv for a private pkg library and project reproducibility
#renv::init()
#renv::activate()

# load packages
install.packages("BiocManager")
library(BiocManager)
options(repos = BiocManager::repositories()) # Bioconductor repo added to renv

# renv::install("tidyverse")
# renv::install("data.table")
# renv::install("RColorBrewer")
# renv::install("VennDiagram")
# renv::install("UpSetR")
# renv::install("patchwork")
# renv::install("viridis")
# renv::install("wesanderson")


#renv::install("stringi")

library(data.table)
library(tinytex)
library(knitr)
library(stringi)
library(UpSetR)
library(tidyverse)
library(RColorBrewer)
library(VennDiagram)
library(grid)
library(patchwork)
library(viridis)
library(wesanderson)

```



### Shared and private SVs in all species ###

```{r upset plot BFF and Eu.polecat SVs, echo=FALSE}


png(file = "UpSetR_mustelids_noBND.png", width = 1800, height = 1500, res = 300)

survivor_matrix <- read.table("input_data/mustelids_manta_wham_lumpy_overlap.txt")
SV_info <- read.csv("input_data/mustelids_manta_wham_lumpy_SV_info.txt", header = F, sep = "\t")
colnames(survivor_matrix) <- c("BFF_REF","SB8055","SB7462","EU_POLECAT") 
                           

#survivor_matrix_all_cols <- survivor_matrix %>% mutate(SV_MantaIDs = mantaIDs) %>% select(SV_MantaIDs, everything())

upset(
      survivor_matrix, 
      nsets = 4, nintersects = 16, 
      mb.ratio = c(0.7, 0.3), matrix.color = "brown1", 
      sets.bar.color = "cornflowerblue", order.by = "freq", 
      keep.order = FALSE, set_size.show = F, 
      text.scale = 1.0, point.size = 2.5, line.size = 1,
      
      )

dev.off()

```


### Distribution of SVs along reference chromosomes ###

```{r SV distribution on chromosomes}

# import list of ordered dom. fer reference chromosomes
ordered_scf_dom_fer <- read.table("input_data/scf_list.tsv") %>% 
                       rename(Scaffolds = "V1") %>%
                       mutate(Chromosomes = c(seq(1:19),"X")) %>%
                       select(Chromosomes, everything())

```

#### BFFs ####

```{r prep BFF stats}

# BFF shared SVs per chromosome stats
bff_chr_sv_distro <- read.table("input_data/BFFs_shared_SVs.stats_CHR", header = T) %>%
                     select(.,-TRA) %>%
                     rename(Scaffolds = "Chr")

bff_chr_ordered <- left_join(ordered_scf_dom_fer, bff_chr_sv_distro, by = "Scaffolds") %>%
                   mutate(Total_SVs = rowSums(.[3:6]))


```



```{r plot BFFs SVs on dom.fer. chromosomes}

# remove unused cols
bff_chr_ordered_long <- bff_chr_ordered %>%
        select(-Scaffolds, -INS, -Total_SVs) %>%
        pivot_longer(!Chromosomes, names_to = "SV_type", values_to = "Count") 
        # extend table into longer format (transpose SV types per Chr)
 
# plot SV distro on chr
png(file = "BFF_shared_SVs.png", width = 1000, height = 800, res = 200)

bff_chr_ordered_long %>%
        mutate(Chromosomes = factor(Chromosomes, levels = unique(Chromosomes))) %>%
      # to keep dataframe order instead of alphabetical, refactor col
        ggplot(aes(x = factor(Chromosomes), y = Count , fill = factor(SV_type))) +
            geom_bar(stat = "identity", position = "dodge") +
            scale_fill_manual(values = c("cornflowerblue", "brown1", "darkgoldenrod1"), name = "SV type") +
            labs(title = "Shared SV distribution on chromosomes",
                 x = "Chromosomes", y = "Number of SVs") +
            theme_light() +
            theme(axis.title.x = element_text(size = 10, vjust = -0.90),
                  axis.title.y = element_text(size = 10, vjust = 1.90),
                  legend.margin = margin(0.5, 0.5, 0.5,0.5)) +
            theme(plot.title = element_text(size = 12,
                  margin = margin(0, 0, 6, 0))) +
            # facet_grid(. ~ SV_type, scales= "free") +
            facet_wrap(. ~ SV_type, ncol=1, scales= "free") # allow diff. scales in plot, y-axis in this case 

dev.off()  


```

#### European polecat ####

```{r prep Eu_polecat stats}

# Eu. polecat
eu_polecat_chr_sv_distro <- read.table("input_data/eu_polecat_only_SVs.stats_CHR", header = T) %>%
                     select(.,-TRA) %>%
                     rename(Scaffolds = "Chr")


eu_polecat_chr_ordered <- left_join(ordered_scf_dom_fer, eu_polecat_chr_sv_distro, by = "Scaffolds") %>%
                   mutate(Total_SVs = rowSums(.[3:6]))

```



```{r plot Eu_polecat SVs on dom.fer. chromosomes}

# remove unused cols
eu_polecat_chr_ordered_long <- eu_polecat_chr_ordered %>%
        select(-Scaffolds, -Total_SVs) %>%
        pivot_longer(!Chromosomes, names_to = "SV_type", values_to = "Count") 
        # extend table into longer format (transpose SV types per Chr)
 
# plot SV distro on chr
png(file = "eu_polecat_SVs.png", width = 1100, height = 950, res = 200)

eu_polecat_chr_ordered_long %>%
        mutate(Chromosomes = factor(Chromosomes, levels = unique(Chromosomes))) %>%
        mutate(SV_type = factor(SV_type, levels = unique(SV_type))) %>%
        # to keep dataframe order instead of alphabetical, refactor cols
        ggplot(aes(x = factor(Chromosomes), y = Count , fill = factor(SV_type))) +
            geom_bar(stat = "identity", position = "dodge") +
            scale_fill_manual(values = c("cornflowerblue", "brown1", "darkgoldenrod1", "darkolivegreen3"), name = "SV type") +
            labs(title = "European polecat SV distribution on chromosomes",
                 x = "Chromosomes", y = "Number of SVs") +
            theme_light() +
            theme(axis.title.x = element_text(size = 10, vjust = -0.90),
                  axis.title.y = element_text(size = 10, vjust = 1.90),
                  legend.margin = margin(0.5, 0.5, 0.5,0.5)) +
            theme(plot.title = element_text(size = 12,
                  margin = margin(0, 0, 6, 0))) +
            # facet_grid(. ~ SV_type, scales= "free") +
            facet_wrap(. ~ SV_type, ncol=1, scales= "free") # allow diff. scales in plot, y-axis in this case

dev.off()  


```

### SV size distribution ###

```{r size bins for each SV type}

# prep stats on SV sizes in BFFs

bff_sv_size_distro <- read.table("input_data/BFFs_shared_SVs.stats", header = T) %>%
                      select(-INS, -TRA, -UNK) %>%
                      slice(-1) %>% # remove 1st row
                      rename_all(toupper) %>%
                      pivot_longer(!LEN, names_to = "SV_type", values_to = "Count") 
                      # extend table into longer format 
```



```{r plot SV size distro}

png(file = "BFF_SV_size_distro.png", width = 1000, height = 800, res = 200)

bff_sv_size_distro %>%
        mutate(LEN = factor(LEN, levels = unique(LEN))) %>%
        ggplot(aes(x = factor(LEN), y = Count , fill = factor(SV_type))) +
            geom_bar(stat = "identity", position = "dodge") +
            scale_fill_manual(values = c("cornflowerblue", "brown1", "darkgoldenrod1"), name = "SV type") +
            labs(title = "SV size distribution in black-footed ferrets",
                 x = "SV size bins", y = "Number of SVs") +
            theme_light() +
            theme(axis.title.x = element_text(size = 10, vjust = -0.90),
                  axis.title.y = element_text(size = 10, vjust = 1.90),
                  legend.margin = margin(0.5, 0.5, 0.5,0.5)) +
            theme(plot.title = element_text(size = 12,
                  margin = margin(0, 0, 6, 0))) #+
            # facet_grid(. ~ SV_type, scales= "free") +
            #facet_wrap(. ~ SV_type, ncol=1, scales= "free") # allow diff. scales in plot, y-axis in this case 

dev.off()  

```

```{r size bins for each SV type}
# prep stats on SV sizes in Eu.polecat

eu_polecat_sv_size_distro <- read.table("input_data/eu_polecat_only_SVs.stats", header = T) %>%
                      select(-TRA, -UNK) %>%
                      slice(-1) %>% # remove 1st row
                      rename_all(toupper) %>%
                      pivot_longer(!LEN, names_to = "SV_type", values_to = "Count") 
                      # extend table into longer format 
```



```{r plot SV size distro}

png(file = "eu_polecat_SV_size_distro.png", width = 1000, height = 800, res = 200)

eu_polecat_sv_size_distro %>%
        mutate(LEN = factor(LEN, levels = unique(LEN))) %>%
        mutate(SV_type = factor(SV_type, levels = unique(SV_type))) %>%
        ggplot(aes(x = factor(LEN), y = Count , fill = factor(SV_type))) +
            geom_bar(stat = "identity", position = "dodge") +
            scale_fill_manual(values = c("cornflowerblue", "brown1", "darkgoldenrod1", "darkolivegreen3"), name = "SV type") +
            labs(title = "SV size distribution in European polecat",
                 x = "SV size bins", y = "Number of SVs") +
            theme_light() +
            theme(axis.title.x = element_text(size = 10, vjust = -0.90),
                  axis.title.y = element_text(size = 10, vjust = 1.90),
                  legend.margin = margin(0.5, 0.5, 0.5,0.5)) +
            theme(plot.title = element_text(size = 12,
                  margin = margin(0, 0, 6, 0))) #+
            # facet_grid(. ~ SV_type, scales= "free") +
            #facet_wrap(. ~ SV_type, ncol=1, scales= "free") # allow diff. scales in plot, y-axis in this case 

dev.off()  

```




### Manhatan plot for each SV type along chromosomes for each individual






### Overlap of detected SVs among all 3 BFF individuals ###

```{r Venn diagram BFF Manta_Wham_Lumpy SV overlaps}

# load overlap table for BFF individuals
bff_survivor_matrix <- read.table("input_data/BFFs_manta_wham_lumpy_overlap.txt")
colnames(bff_survivor_matrix) <- c("BFF_REF","SB8055","SB7462") 
diagram_colors <- brewer.pal(3, "Set1")

venn.diagram(
        list(
          BFF_REF=which(bff_survivor_matrix[,1]==1), 
          SB8055=which(bff_survivor_matrix[,2]==1), 
          SB7462=which(bff_survivor_matrix[,3]==1)
        ), 
        # colors, lines
        fill = diagram_colors, 
        lwd = 2, 
        lty ='blank', 
        alpha = c(0.5, 0.5, 0.5), 
        # numbers
        cex = 0.7, 
        fontface = "bold",  
        fontfamily = "sans",
        # category names
        cat.cex = 0.7,
        cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-20, 20, 125),
        cat.dist = c(0.055, 0.055, 0.075),
        cat.fontfamily = "sans",
        rotation = 1,
        # output features
        imagetype="png" ,
        height = 750 , 
        width = 750 , 
        resolution = 300,
        compression = "lzw",
        filename = "BFFs_overlap.png"
)


```


### Insert size distribution for each individual ###


### Coverage distibution along chromosomes for each individual ###




